---
title: "ADA Project"
author: "Wenqing Zha"
date: '2022-12-04'
output: html_document
---
# Install packages
```{r}
pacman::p_load(haven, foreign, odds.n.ends, blorr, lmtest, car,  broom, tidyverse, jtools) 
```

# import dataset
```{r}
Demo1314 <- read_xpt(file = "DEMO_H.xpt")
Demo0910 <- read_xpt(file = "DEMO_F.xpt")

Fracture1314 <- read_xpt(file = "OSQ_H.xpt")
Fracture0910 <- read_xpt(file = "OSQ_F.xpt")

Outdoor1314 <- read_xpt(file = "DEQ_H.xpt")
Outdoor0910 <- read_xpt(file = "DEQ_F.xpt")
```

# Select variables we need
```{r}
Demo1314 <- Demo1314 %>%
  select(c("SEQN", "RIAGENDR", "RIDAGEYR", "RIDRETH1"))
names(Demo1314) <- c("Participant", "Gender", "Age", "Race")

Demo0910 <- Demo0910 %>%
  select(c("SEQN", "RIAGENDR", "RIDAGEYR", "RIDRETH1"))
names(Demo0910) <- c("Participant", "Gender", "Age", "Race")


Fracture1314 <- Fracture1314 %>%
  select(c("SEQN", "OSQ010A", "OSQ010B", "OSQ010C"))
names(Fracture1314) <- c("Participant", "hip_frac", "wrist_frac", "spine_frac")

Fracture0910 <- Fracture0910 %>%
  select(c("SEQN", "OSQ010A", "OSQ010B", "OSQ010C"))
names(Fracture0910) <- c("Participant", "hip_frac", "wrist_frac", "spine_frac")


Outdoor1314 <- Outdoor1314 %>%
  select(c("SEQN", "DED120", "DED125"))
names(Outdoor1314) <- c("Participant", "Outdoor_workday", "Outdoor_non_workday")

Outdoor0910 <- Outdoor0910 %>%
  select(c("SEQN", "DED120", "DED125"))
names(Outdoor0910) <- c("Participant", "Outdoor_workday", "Outdoor_non_workday")
```

# Merge 2009-2010 and 2013-2014 data
```{r}
Demo <- rbind(Demo0910, Demo1314)

Fracture <- rbind(Fracture0910, Fracture1314)

Outdoor <- rbind(Outdoor0910, Outdoor1314)

NHANES <- full_join(Demo, Fracture, by="Participant")
NHANES <- full_join(NHANES, Outdoor, by="Participant")
```

# select participants aged 45 or older
```{r}
NHANES <- filter(NHANES, Age >= 45)
```

```{r}
# Gender
table(NHANES$Gender)

NHANES <- NHANES %>%
  mutate(Gender_binary = case_when(Gender %in% c(2) ~ 0,
                                     Gender %in% c(1) ~ 1),
         Gender_binary = factor(Gender_binary, 0:1, c("Female", "Male")))

table(NHANES$Gender_binary, useNA="always")

# Race
table(NHANES$Race)

NHANES <- NHANES %>%
  mutate(Race_recode = factor(Race, 1:5, c("Mexican American", "Other Hispanic", "Non-Hispanic White", "Non-Hispanic Black", "Other race")))

table(NHANES$Race_recode, useNA="always")

# Fracture
NHANES <- NHANES %>%
  mutate(NHANES, Fracture_recode = ifelse(hip_frac == 1 | wrist_frac == 1 | spine_frac == 1, 1, 0),
         Fracture_recode = factor(Fracture_recode, 0:1, c("no fracture", "fracture")))

table(NHANES$Fracture_recode, useNA="always")

# Outdoor
NHANES$Outdoor <- NHANES$Outdoor_workday + NHANES$Outdoor_non_workday
```

## Run a univariable (unadjusted) logistic model for outdoor as a continuous variable
```{r}
model1<- glm(Fracture_recode ~ Outdoor, data=NHANES, family="binomial")
summary(model1) # get log results

## interpretation: For every one unit in bmi, the log odds of diabetes increase 0.090185

odds.n.ends(model1) # get OR results (and more!)
```
# Test the linearity assumption for Outdoor using the Box Tidwell test
```{r}
NHANES <- NHANES %>%
  mutate(Outdoor.times.logOutdoor = Outdoor * log(Outdoor)) # create term to test linearity

model2 <- glm(Fracture_recode ~ Outdoor + Outdoor.times.logOutdoor, data=NHANES, family="binomial") 

summary(model2)

# a NON-SIGNIFICANT (p > .05) coefficient means the assumption IS NOT VIOLATED. It is linearity.
```
## Adjust for age, gender, race category
```{r}
model3 <- glm(Fracture_recode ~ Outdoor + Gender + Age + Race, data=NHANES, family="binomial")
summary(model3)

odds.n.ends(model3)

compareCoefs(model1, model3)
```
# Look at assumptions of multicollinearity using the vif function from the car package
```{r}
vif(model3)

# VIF = 1, no correlation between the independent variable and the other variables
```
# Look for influential observations using Cook's distance. First we will plot
```{r}
# Plot Cooks's Distance
plot(model3, which = 4, id.n = 3, col="red") 

```
## Next we will get Cook's D calculations for each obervation and some other stats using the augment function  from the Broom package.
```{r}
model3.data <- augment(model3) %>%  
  mutate(index = 1:n()) 

head(model3.data)
```
## Identify and exclude observations with a Cook's Distance greater than a cutoff, use 3 x the mean Cook's D 
```{r}
cuttoff <- mean(model3.data$.cooksd)*3 # defines cutoff variable

NHANES_out <- model3.data %>% # excludes observations
  filter(.cooksd<cuttoff)

nrow(NHANES_out)/nrow(NHANES) # check percent kept
```
## Model Fracture and Outdoor logistic model with observations excluded
```{r}
model4 <- glm(Fracture_recode ~ Outdoor + Gender + Age + Race, data=NHANES, family="binomial")
summary(model4)
odds.n.ends(model4)
```
# Compare coefficients between models with and without influential observations 
```{r}
compareCoefs(model3, model4) 
```
```{r}
odds.n.ends(model4, predProbPlot=TRUE, rocPlot = TRUE)
```






